---
title: "Biostat 653 Hw1"
author: "David (Daiwei) Zhang"
date: "September 27, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(matrixStats)
require(car)
```

## Problem 6

### (1)

Our model is 
\[
  Y \sim \beta_0 + \beta_1 I(\text{girl}) + \beta_2 I(\text{cur. expo.}) + \beta_3 I(\text{past. expo.}) + \beta_4 I(\text{cur. expo.}) I(\text{girl}) + \beta_5 I(\text{past. expo.}) I(\text{girl}) 
\]
Then the three questions are hypothesis tests with
\begin{enumerate}
\item
\begin{align*}
  H_0 &: \beta_2 = \beta_3 = \beta_4 = \beta_5 = 0 \\
  H_1 &: \text{At least one of these is nonzero.}
\end{align*}
\item
\begin{align*}
  H_0 &: \beta_4 = \beta_5 = 0 \\
  H_1 &: \text{At least one of these is nonzero.}
\end{align*}
\item
\begin{align*}
  H_0 &: \beta_2 < \beta_3 \quad\text{and}\quad \beta_2 + \beta_4 < \beta_3 + \beta_5 \\
  H_1 &: \text{At least one of these is false.}
\end{align*}
\end{enumerate}

We fit the linear model:
```{r}
leadiq <- read.table("~/leadiq.txt")
colnames(leadiq) <- c("id", "expo.cat", "gender", "age", "iq")
leadiq$expo.cat <- factor(leadiq$expo.cat)
leadiq.model <- lm(iq ~ gender + expo.cat + expo.cat * gender, data=leadiq)
summary(leadiq.model)
```

### (2)
We do not need to include age as a covariate, because the investigator does not ask for adjustment for age.
Moreover, if we include age in our model,
```{r}
leadiq.model.withage <- lm(iq ~ age+ gender + expo.cat + expo.cat * gender, data=leadiq)
summary(leadiq.model.withage)
```

We see that age is not a significant predictor at $\alpha = 0.05$.

### (3)
$\beta_0$: The mean IQ of a boy with no lead exposure is 103.696 ($p < 2 \times 10^{-16}$). 

$\beta_1$: The mean IQ of a girl is 2.414 ($p = 0.508$) lower than that of a boy, given that they have no lead exposure. 

$\beta_2$: The mean IQ of a boy with current lead exposure is 6.284 ($p = 0.163$) lower than that with no lead exposure.

$\beta_3$: The mean IQ of a boy with past lead exposure is 10.234 ($p = 0.041$) lower than that with no lead exposure. 

$\beta_4$: The mean difference in IQ from a girl with current lead exposure to that with no lead exposure is 3.57 ($p = 0.655$) lower than that of two boys in such situations. 

$\beta_5$: The mean difference in IQ from a girl with past
lead exposure to that with no lead exposure is 4.06 ($p = 0.601$) higher than that of two boys in such situations. 

### (4)
IQ of boy without lead exposure: $\beta_0 = 103.696$.
IQ of boy with current lead exposure: $\beta_0 + \beta_2 = 97.412$.
IQ of boy with past lead exposure: $\beta_0 + \beta_3 = 93.462$.

### (5)

As seen from the qq plot below, the observed data is close to the straight line, so which supports the linearity assumption. 
However, in the absolute standardized residual plot,
we see that the absolute residuals increase with the fitted values,
so the constant variance assumption is slightly violated.

```{r}
par(mfrow=c(1,2))
qqPlot(leadiq.model, main = "qq plot")
spreadLevelPlot(leadiq.model)
```

## Problem 7

### (1)
```{r}
lead <- read.table("~/lead.txt")
colnames(lead) <- c("id", "week.0", "week.1", "week.4", "week.6")
lead.time <- c(0,1,4,6)
lead.mat <- as.matrix(lead[,-1])
lead.mean <- colMeans(lead.mat)
lead.sd <- colSds(lead.mat)
lead.var <- lead.sd^2
print(lead.mean)
print(lead.sd)
print(lead.var)
```

### (2)
Overall, the lead level decrease sharply from the baseline to the first week, and then increases slowly till the sixth week.
```{r}
matplot(lead.time, t(lead.mat), type="l", lty=2, lwd=1, col=1)
points(lead.time, lead.mean, type="l", lty=1, lwd=5, col=1)
```

###(3)
The two calculations of the variance are identical.
```{r}
lead.cov <- cov(lead.mat)
lead.cor <- cor(lead.mat)
print(lead.cov)
print(lead.cor)
```



