---
title: "Biostat 653 Homework 2"
author: "David (Daiwei) Zhang"
date: "October 11, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(matrixStats)
require(MASS)
rm(list=ls())
set.seed(10)
```

## Problem 2

### (1)
```{r}
N <- 1000
mu <- c(0,0)
sig <- matrix(c(2, 0.5, 0.5, 1), 2, 2)
y <- mvrnorm(N, mu, sig)
head(y)
y <- t(y)
```

###(3)

```{r}
x <- matrix(1,2,1)
sig <- matrix(c(1,0,0,1), 2, 2)
num.iter <- 1e4
for(i in 1:num.iter){
  sig.inv <- solve(sig)
  mu <- 1/N / sum(sig.inv) * sum(colSums(sig.inv) %*% y)
  sig <- (y - mu) %*% t(y - mu) / N
}
mu.var <- 1/sum(sig.inv) / N
print(mu)
print(sig)
print(mu.var)
```

### (4)

Test statistic: $N(1_2\hat{\mu})^T{\hat{\Sigma}}^{-1}(1_2\hat{\mu}) \sim \chi_1^2$.
```{r}
test.stat <- N * sum(sig.inv) * mu^2 
pval <- 1 - pchisq(test.stat, 1)
print(pval)
```
Since the p-value is $0.583196$, we do not rejec the null hypothesis at $\alpha = 0.05$.

### (6)
```{r}
x <- matrix(1,2,1)
w <- matrix(c(1,0,0,1), 2, 2)
num.iter <- 1e4
for(i in 1:num.iter){
  mu <- 1/N / sum(w) * sum(colSums(w) %*% y)
  sig <- (y - mu) %*% t(y - mu) / N
  a <- sum(diag(sig)) / 2
  b <- sig[1,2]
  sig <- matrix(c(a,b,b,a), 2,2)
  w <- solve(sig)
}
mu.var <- 1/sum(sig.inv) / N
print(mu)
print(sig)
print(mu.var)
```

### (7)
We see that $\hat{\mu}$ is close to $\mu$, and the off diagonal elements of $\hat{\sum}$ is close to those of $\sum$.
However, the diagonal elements are the average of the true diagonal elements, which makes sense given the restriction of our model.

### (8)
```{r}
mu.var.model <- 1 / sum(w)/N
mu.var.robust <- (1/N)*(1/sum(w)) * sum((colSums(w)%*%(y-mu))^2) * (1/sum(w))*(1/N)
print(mu.var.model)
print(mu.var.robust)
```
The two variances are close here. However, if they are different, then the robust variance is better to use, since our model is not correct.

## Problem 3

### 5.1.1

```{r}
col.names = c("group", "id", "cho0", "cho6", "cho12", "cho20", "cho24")
cho <- read.table("../Downloads/cholesterol-data.txt", na.strings = ".", col.names = col.names)
head(cho)
```

### 5.1.2
```{r}
means <- c()
vars <- c()
sds <- c()
for(i in 1:2){
    print(i)
    mean.this <- colMeans(cho[cho$group==i,-(1:2)], na.rm = TRUE)
    var.this <- colVars(as.matrix(cho[cho$group==i,-(1:2)]), na.rm = TRUE)
    sd.this <- colSds(as.matrix(cho[cho$group==i,-c(1:2)]), na.rm = TRUE)
    means <- rbind(means, mean.this)
    vars <- rbind(vars, var.this)
    sds <- rbind(sds, sd.this)
}
rownames(means) <- c(1,2)
rownames(vars) <- c(1,2)
rownames(sds) <- c(1,2)
colnames(vars) <- colnames(means)
colnames(sds) <- colnames(means)
print(means)
print(vars)
print(sds)
```

### 5.1.3
```{r}
occasions <- c(0, 6, 12, 20, 24)
matplot(occasions, t(means), type="b")
```

Cholesterol levels increase over time for both groups. 
Compared to that of the placebo group,
the cholesterol level of the high-dose group is lower at baseline and the end but higher in the middle.

### 5.1.4
```{r}
cho.uni <- c()
for(i in 1:5){
    cho.uni.this <- cho[,c(1,2,2+i)]
    colnames(cho.uni.this) <- c("group", "id", "cho")
    cho.uni.this$occasion <- occasions[i]
    cho.uni <- rbind(cho.uni, cho.uni.this)
}
head(cho.uni)
write.table(cho.uni, "biostat653_data_cho.uni", row.names = FALSE, col.names = FALSE)
```

### 5.1.5
See SAS output.
The p-value for the F-statistic for the interaction of group by occasion is $0.0990$,
so we do not reject the null hypothesis that the patterns of change over time is different in the two groups at $\alpha = 0.05$.

### 5.1.6
See the SAS output for the covariance matrix ("Estimated R Matrix for ID 1") and the correlation matrix ("Estimated R Correlation Matrix for ID 1")

### 5.1.7
The model is
\begin{align*}
  cholesterol_i & = \beta_0 \\
  & + \beta_1 I(occasion=6) + \beta_2 I(occasion=12) \\
  & + \beta_3 I(occasion=20) + \beta_4 I(occasion=24) \\
  & + \beta_5 I(group=1) \\
  & + \beta_6 I(group=1)I(occasion=6) + \beta_7 I(group=1)I(occasion=12) \\
  & + \beta_8 I(group=1)I(occasion=20) + \beta_9 I(group=1)I(occasion=24)
\end{align*}

### 5.1.8
\[
L = 
\begin{bmatrix}
0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 
\end{bmatrix}
\]

### 5.1.9
We can simply add up the effects to get the means.
For example,
\begin{align*}
  E[group=2, occasion=0] & = 235.93 \\
  E[group=2, occasion=6] & = 235.93 + 7.2439 = 243.1739 \\
  E[group=1, occasion=0] & = 235.93 + (-9.9107) = 226.0193 \\
  E[group=1, occasion=6] & = 235.93 + (-9.9107) + 7.2439 + 12.2722 = 245.5354
\end{align*}
This gives the same result from the group means calculated earlier.

### 5.1.10
Interpretations: 

$\beta_0$: The mean cholesterol level for an individual in the placebo group at baseline.

$\beta_1$: The mean difference in cholesterol level for an individual in the placebo group from baseline to 6 months after.

$\beta_2$ to $\beta_4$: Similar to $\beta_1$.

$\beta_6$: The mean difference in cholesterol level at baseline from an individual in the placebo group to an individual in the high-dose group.

$\beta_7$: The mean group effect on the difference in cholestorol level from baseline to 6 months after.

$\beta_8$ and $\beta_9$: Similar to $\beta_7$.
