\documentclass[12pt]{article}
\usepackage{float}
% \usepackage[nolists]{endfloat}
\usepackage{epsfig} \usepackage[margin=1.5in]{geometry}
\usepackage[semicolon,authoryear]{natbib} \bibliographystyle{natbib}
\usepackage{amsmath}
\usepackage{cleveref}

\begin{document}
\setlength{\textheight}{625pt} \setlength{\baselineskip}{23pt}

\title{Analysis of Error in the 2016 Presidential Election Polls}
\author{David (Daiwei) Zhang}
\date{Apr 26, 2017}
\maketitle

\noindent

\section{Introduction} \label{sec:intro}

The 2016 presidential election generated many interesting datasets and statistical problems.
Most polls have underestimated Trump's performance,
including state polls, national polls, forecasts, and exit polls (\cite{bialik}).
Exports have suggested different survey factors that caused this result,
especially the nonresponse bias and low turnout rate (\cite{stein}).
In this report, we present an analysis of the election poll and result datasets in order to better understand the errors in the polls.

\section{Datasets} \label{sec:data}

The analysis uses mainly two datasets, one for the polls and the other for the election results.
The poll dataset is obtained from FiveThirtyEight (\cite{pollData}).
It consists of state level and national poll results conducted by media, universities, and other institutes.
Beside the proportion of votes won by each candidate, the dataset also contains information such as the sample size and the date of the poll.
The election result is contained in an unpublished dataset compiled by Emil O. W. Kirkegaard from the New York Times and a study of inequality in the U. S. (\cite{kirkegaard}).
It contains county-level election result as well as more than a hundred variables about each county's demography, economy, and other information.
Both datasets are available to the public.

\section{Analysis} \label{sec:analy}

We are interested in how the poll's deviation from the election result is associated with other factors.

Thus we conduct a linear regression analysis of the absolute 


\section{Discussion} \label{sec:discuss}





\newpage

\begin{thebibliography}{}

\bibitem[Bialik \& Enten(2016)]{bialik} Bialik, C. and Enten, L. The Polls Missed Trump. We Asked Pollsters Why. FiveThirtyEight (2016).

\bibitem[Silver (2016)]{pollData} Silver, N. Presidential General Polls, 2016. FiveThirtyEight (2016). Retrieved from https://projects.fivethirtyeight.com/2016-election-forecast.

\bibitem[Kirkegaard (2016)]{voteData} Kirkegaard, E. O. USA County Data. (2016). Unpublished dataset. Retrieved from https://github.com/Deleetdk/USA.county.data.

\bibitem[Kirkegaard (2016)]{kirkegaard} Kirkegaard, E. O. Inequality across US counties: an S factor analysis. (2016).

\bibitem[Stein(2016)]{stein} Stein, J.. ``7 experts try to explain how the polls missed Donald Trump’s victory''. Vox (2016).



\end{thebibliography}

\end{document}
