---
title: "Survmeth 895 Homework 3"
author: "David (Daiwei) Zhang"
date: "February 1, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 3

The posterior distribution of the population mean is 
\begin{align*}
  \bar{Y} | y_{\text{inc}} 
  & \sim
  (\bar{Y} | \mu, \sigma^2, y_{\text{inc}})
  (\mu | \sigma^2, y_{\text{inc}})
  (\sigma^2 | y_{\text{inc}})
  \\ & \sim
  [(n-1)s^2 \chi^2_{n-1}]
  [N(\bar{y}_{\text{inc}}, \sigma^2/n)]
  [N(\mu, \sigma^2 / k)]
\end{align*}
We can use simulation to get more concrete results.

```{r}
a = c(60, 72, 78, 92, 78, 57, 62, 72, 88, 71)
b = c(68, 90, 88, 88, 68, 72, 72, 92, 50, 99)
y = cbind(a, b)
sampsize = length(a)
ybar = c(mean(a), mean(b))
ssquare = c(sd(a), sd(b))
k = 120
nsimul = 10000
result = matrix(0,nsimul,8)
for (i in 1:nsimul){
  tmp=rnorm(sampsize-1)
  chisq=sum(tmp*tmp)
  sigmasq=(sampsize-1)*ssquare/chisq;
  mu=ybar+sqrt(sigmasq/
                 sampsize)*rnorm(1)
  ybark=mu+sqrt(sigmasq/k)*rnorm(1)
  yind=mu+sqrt(sigmasq)*rnorm(1)
  result[i,] = c(sigmasq, mu, ybark, yind)
}
par(mfrow=c(3,2))
hist(result[,1], xlab="ssquare.a")
hist(result[,2], xlab="ssquare.b")
hist(result[,3], xlab="mu.a")
hist(result[,4], xlab="mu.b")
hist(result[,5], xlab="ybark.a")
hist(result[,6], xlab="ybark.b")
```

### (a)

We calculate the $95\%$ credibility intervals for the population mean for firm A and firm B

```{r}
## Firm A
quantile(result[,5], c(0.025, 0.975))
## Firm B
quantile(result[,6], c(0.025, 0.975))
```

### (b)

The proficient rate is

```{r}
## Firm A
sum(result[,7] >= 75) / nsimul
## Firm B
sum(result[,8] >= 75) / nsimul
```
### (c)

Based on the credibility intervals and proficiency rates, Firm B provides significantly better training than Firm A.